On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-02 main.tex

Let $\A$ be an $n \times n$ matrix of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in \mathbb{Q}(\A)[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$ Recall from Section 4.1.4 that $f^\A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2),$ and $\textbf{H}_i'(3).$ Throughout this section, for simplicity let $\I^\A$ denote $\I(\pi_i,V^\A)$ and let $W^\A$ denote $W(\pi_i,V^\A).$ To increase readability and simplify notation, we often do not specify $i$, as it is fixed.   

@@

Let $f \in \mathbb{Q}[X_1,\hdots,X_n]$ be squarefree with total degree $D$ and $V(f)$ smooth. For $i \in \{1,\hdots,n\}$, recall $\textbf{H}_i'(4):$ \textit{$W(\pi_i,V)$ is in Noether position for $\pi_{i-1}$}, which is equivalent to having the extension  

@@

@@

then $A$ cancels none of the denominators of $a_{j}$ or $b_{j,l}$. Define 

@@

\Xi_j:=V(h_j), 

@@

where $h_j$ is the product of the denominators of the coefficients of $a_j$ and $b_{j,l}$, so that when $A \not \in \Xi_j$ then $A$ cancels none of the denominators $a_{j}$ or $b_{j,l}$ which provide the integral dependence relation. For each $j$, we want a bound on the degree of $\Xi_j$. We can obtain bounds by applying an effective Nullstellensatz theorem. However, the theorem requires $P_{j}$ to be a polynomial. We can obtain a polynomial from $P_{j}$ by clearing denominators. We can multiply $P_{j}$ by $d_j \in \Q[\A_{i,j}]$ equal to the least common multiple of the denominators of $P_{j}.$ Put 

@@

\overline{P_{j}} := dP_{j}\in \Q[\A][\textit{\textbf{X}}],

@@

\deg d \leq \deg \pjb \leq (D+1)^{n-i+1}.

@@

A_l(\A_{i,j},\textit{\textbf{X}},1/\pjb) &= \sum_{i=1}^{n_l} A_{l,i}(\A_{i,j},\Xb,1/\pjb)\\
&= \sum_{i=1}^{n_l} \left(\frac{1}{P_j^{k_l}}\right)\At_{l,i}(\A_{i,j},\textit{\textbf{X}}), ~k_l \in \mathbb{Z}.

@@

Now let $k := \max\{k_l\}$ so that 

@@

    \pjb^k \alpha &= \pjb^k\left(\sum_{l=1}^{n-i+1} A_l(\A_{i,j},\textit{\textbf{X}},1/\pjb)g_l\right)\\
    &= \sum_{l=1}^{n-i+1}\left( \sum_{i=1}^{n_l}\left(\pjb^{k-k_{n_l}}\right)\At_l(\A_{i,j},\Xb) \right)g_l \in \Q[\A_{i,j}][\Xb],

@@

which shows that $\alpha$ is the only remaining denominator. We know that $\alpha$ is not zero, we know the leading coefficient of $\overline{P_j}$ is not zero, we have an upper bound for the degree of $\alpha$, and we know $\deg d \leq \deg P_j = (D+1)^{n-i+1}$, where $\pjb=dP_j.$ By making $\pjb$ monic, the integrality of 
\begin{align*}\mathbb{Q}(\frak{A}_{i,j})[\textbf{X}_{\leq i-1}]\rightarrow\mathbb{Q}(\frak{A}_{i,j})[\textbf{X}]/\I_i^{\A}

@@

holds with $\pjb(X_j)\in \I_i^{\A}.$

@@

\subsection{Proof of Theorem 5.1?}

@@

There is a Zariski-open subset $\Gamma \subset \GL(n,\C)$ such that for $A \in \Gamma$, the entries of $A$ cancel none of the denominators of the coefficients of the generators of $\I(\pi_i,V^{A}).$ Thus, we set

@@

with $h \in \Q[\A_{i,j}]$ defined as the least common multiple of the denominators of the generators of $\I^\A.$ Since $\pjb^k \in \I^\A$ and since
\pjb^k = \frac{1}{\alpha}\left( Q(\A_{i,j},\Xb) \right), Q \in \Q[\A_{i,j}][\Xb],
it follows that 
\deg \Xi = \deg h \leq \deg \alpha \leq D^{\frac{3}{2} n \ln n+1}.
Furthermore, since $\pjb^k \in \I^\A,$ it follows from by proposition 5.2 that, if $A \in \C^{n^2} - \Xi$ then the extension  
\C[\textbf{X}_{\leq i-1}]\rightarrow\C[\textbf{X}]/\frak{I}^\A
is integral, and thus $V(\I^A)=W^\A$ is in Noether position for $\pi_{i-1}.$ Therefore, for each $i \in\{1,\hdots,n\}, f^A$ satisfies $\textbf{H}_i'(4)$.

>>>>

%% <<<<<<< HEAD
%% Let $\A$ be an $n \times n$ matrix of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in \mathbb{Q}(\A)[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$ Recall from Section 4.1.4 that $f^\A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2),$ and $\textbf{H}_i'(3).$ Throughout this section, for simplicity let $\I^\A$ denote $\I(\pi_i,V^\A)$ and let $W^\A$ denote $W(\pi_i,V^\A).$ To increase readability and simplify notation, we often do not specify $i$, as it is fixed.   
%% \subsection{Sketch of the proof}
%% Let $f \in \mathbb{Q}[X_1,\hdots,X_n]$ be squarefree with total degree $D$ and $V(f)$ smooth. For $i \in \{1,\hdots,n\}$, recall $\textbf{H}_i'(4):$ \textit{$W(\pi_i,V)$ is in Noether position for $\pi_{i-1}$}, which is equivalent to having the extension  
%% =======
Let $\A$ be an $n \times n$ matrix of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in \mathbb{Q}(\A)[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$ Recall from Section 4.1.4 that $f^\A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Throughout this section, to simplify notation, let $\I^\A$ denote $\I(\pi_i,V^\A)$ and let $W^\A$ denote $W(\pi_i,V^\A).$ Furthermore, throughout this section, consider that $i \in \{1,\hdots,n\}$ is fixed. We will avoid writing $i$ where the context is clear. 

@@

Let $f \in \mathbb{Q}[X_1,\hdots,X_n]$ have total degree $D$. Assume $f$ satisfies \textbf{H}, so that $f$ is squarefree with $V(f)$ smooth. Recall $\textbf{H}_i^{'}(7):$ \textit{$W(\pi_i,V)$ is in Noether position for $\pi_{i-1}$}, which is equivalent to having the extension  
%% >>>>>>> f9eb2445ca6e153d1e49c3d2767097d331dbcbee

@@

Section 8 is devoted to proving Theorem 8.1. It has already been shown in \cite{EMP} that $f^A$ satisfies $\textbf{H}_i^{'}$ for a generic change of variables. The new contribution is in upper bounding the degree of $\Delta_3.$

@@

then $A$ cancels none of the denominators of $a_{j}$ or $b_{j,l}$. Define $h_j \in \Q[\A]$ to be the product of the denominators of the coefficients of $a_j$ and $b_{j,l}$. Hence, if $A \in \Q^{n^2}-V(h_j)$ then $A$ cancels none of the denominators of $a_{j}$ or $b_{j,l},$ which are providing the integral dependence relation. Now we define 

@@

\Xi_i := V(h_i,\hdots,h_n), 

@@

and we obtain a bound on the degree of $\Xi_i$ by applying Bezout's theorem. Hence, if $A\in \Q^{n^2}-\Xi_i$ then

@@

P_j(X_j) \in \I^A,  
for each $j \in \{i,\hdots,n\},$ and therefore the extension 
\mathbb{C}[{\Xb}_{\leq i-1}]\rightarrow\mathbb{C}[\Xb]/\I(\pi_i,V^{A})
is integral. We can obtain bounds for $h_j$ by applying an effective Nullstellensatz theorem. However, the theorem requires $P_{j}$ to be a polynomial. We can obtain a polynomial from $P_{j}$ by clearing denominators. We can multiply $P_{j}$ by the least common multiple of its denominators, which we will denote by $d_j \in \Q[\A_{i,j}]$. Put 
\overline{P_{j}} := d_jP_{j}\in \Q[\A][\textit{\textbf{X}}],

@@

\deg d_j \leq \deg \pjb \leq \Dt_j.

@@

A_{j,l}(\A,\textit{\textbf{X}},1/\pjb) 
&= \sum_{i=1}^{n_l} A_{j,l,i}(\A,\Xb,1/\pjb)\\
&= \sum_{i=1}^{n_l} \left(\frac{1}{P_j^{k_i}}\right)\At_{j,l,i}(\A,\textit{\textbf{X}}), ~k_i \in \mathbb{Z}.

@@

Let $K := \max\{k_l\}$ so that 

@@

    \pjb^K \alpha_j &= \pjb^K\left(\sum_{l=1}^{n-i+1} A_{j,l}(\A,\textit{\textbf{X}},1/\pjb)g_l\right)\\
    &= \sum_{l=1}^{n-i+1}\left( \sum_{i=1}^{n_l}\left(\pjb^{K-k_{i}}\right)\At_{j,l,i}(\A,\Xb) \right)g_l \in \Q[\A][\Xb].

@@

which shows that $\alpha_j$ is the only remaining denominator whose degree we need to bound other than $d_j$. Note as well that
\pjb^K \in \I^{\A}.
Therefore 
\deg h_j &\leq \left(\deg \alpha_j\right)\left(\deg d_j\right)\\
    &\leq \left(D^n\left(\Dt_j + n\left(\Dt_j+1\right) \right)\right)\Dt_j\\
    &=D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right).

@@

@@

\subsection{Proof of Theorem 8.1}

@@

We can apply B\'ezout and obtain
    \deg \Xi_i = \deg V(h_i,\hdots,h_n)
    &\leq \prod_{j=i}^n \deg h_j\\
    &\leq  \prod_{j=i}^n \left( D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) \right) \\
    &=  \left( D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) \right)^{n-i+1}\\
    &\leq  \left( D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) \right)^{n}.
And therefore, by another application of B\'ezout, 
    \deg& \Delta_3 \leq \prod_{i=1}^n \deg \Xi_i\\
    &\leq \left( D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) \right)^{n^2},
    &(=D^{O(n^3 \log n)}\textrm{?})
where if $A \in \C^{n^2} - \Delta_3$ then, for each $i \in \{1,\hdots,n\},$ the extension 

@@

%% <<<<<<< HEAD
%% is integral, and thus $V(\I^A)=W^\A$ is in Noether position for $\pi_{i-1}.$ Therefore, for each $i \in\{1,\hdots,n\}, f^A$ satisfies $\textbf{H}_i'(4)$.
%% =======
%% >>>>>>> f9eb2445ca6e153d1e49c3d2767097d331dbcbee
